---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r loadlibraries, echo=FALSE, results='hide'}
library("dplyr")
library(ggplot2)
```

## Loading and preprocessing the data

```{r loaddata}
setwd("~/DataScience/Reproducible Research")
activitydata <- read.csv(unz("activitydata.zip", "activity.csv"), sep=",")
activitydata$date <- as.Date(activitydata$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r meansteps}
stepsperday <- activitydata %>%   group_by(date) %>%   summarise(steps = sum(steps))
hist(stepsperday$steps, main="Histogram of total steps per day" , ylim=c(0,40))
meanstepsperday <- round(mean(stepsperday$steps, na.rm=TRUE),2)
medianstepsperday <- round(median(stepsperday$steps, na.rm=TRUE),0)
meanstepsperday
medianstepsperday
```
The mean of total number of steps taken per day is `r meanstepsperday`  
The median of total number of steps taken per day is `r medianstepsperday`


## What is the average daily activity pattern?

```{r dailyactivity}
avgstepsperinterval <- activitydata %>%   group_by(interval) %>%   summarise(steps = mean(steps, na.rm=TRUE))
ggplot(avgstepsperinterval, aes(x=interval, y=steps)) + geom_line() +  labs(title ="Average daily activity pattern", x = "Time", y = "Average Steps") 
maxinterval <- avgstepsperinterval[avgstepsperinterval$steps == max(avgstepsperinterval$steps),1]
maxsteps <- avgstepsperinterval[avgstepsperinterval$steps == max(avgstepsperinterval$steps),2]
maxinterval
maxsteps
```
On average across all the days the maximum number of steps is taken in the 5-minute interval at `r maxinterval`
The average number of steps taken in this interval is `r maxsteps`

## Imputing missing values

``` {r missingvalues}
totalmissing <- sum(is.na(activitydata))
totalmissing
activitydatacompleted <- activitydata

activitydatacompleted[is.na(activitydatacompleted$steps),"steps"] <- round(avgstepsperinterval[match(activitydatacompleted$interval, avgstepsperinterval$interval), 2])

stepsperdaycompleted <- activitydatacompleted %>%   group_by(date) %>%   summarise(steps = sum(steps))
hist(stepsperdaycompleted$steps, main="Histogram of total steps per day (completed)" , ylim=c(0,40))
hist(stepsperday$steps, main="Histogram of total steps per day" , ylim=c(0,40))

meanstepsperdaycomplete <- mean(stepsperdaycompleted$steps, na.rm=TRUE)
medianstepsperdaycomplete <- median(stepsperdaycompleted$steps, na.rm=TRUE)
meanstepsperdaycomplete
medianstepsperdaycomplete
```
Total number of missing values in the dataset is `r totalmissing`
The mean of total number of steps taken per day (completed dataset) is `r meanstepsperdaycomplete`
The median of total number of steps taken per day (completed dataset) is `r medianstepsperdaycomplete`



## Are there differences in activity patterns between weekdays and weekends?

``` {r weekendsactivity}
activitydata$weekday = "" 
activitydata$weekday <- weekdays(activitydata$date)

activitydata[activitydata$weekday!=c("Saturday", "Sunday"), "weekday"] <- "weekday"
activitydata[activitydata$weekday==c("Saturday", "Sunday"), "weekday"] <- "weekend"
activitydata$weekday <- as.factor(activitydata$weekday)

avgstepsperinterval1 <- activitydata %>%   group_by(interval,weekday) %>%   summarise(steps = mean(steps, na.rm=TRUE))

p<- ggplot(avgstepsperinterval1, aes(x=interval, y=steps)) + geom_line() + facet_grid(weekday~.)
p <- p + labs(title ="Average daily activity pattern", x = "Time", y = "Average Steps") 
p
```
Average number of steps during weekends are generally lower except between 0730 and 1000 hours. 



